Evaluation of classical correlation functions from 2/3D images on CPU and GPU architectures: introducing CorrelationFunctions.jl

### 2021-07-13 paper-correlations.tex
Minor edits + few words about cross-correlation functions

\section{Overview of supported functions and modules.}
\label{oversec}
\verb+CorrelationFunctions.jl+ is capable to compute the following correlation
functions:
\begin{itemize}
\item Lineal-path function $L_2$.
\item Two-point function $S_2$.
\item Cluster function $C_2$.
\item Surface-surface function $F_{ss}$.
\item Surface-void function $F_{sv}$.
\item Pore size function $P$.
\item Chord length function $p$.
\end{itemize}

@@

\subsection{Lineal-path function}
Lineal-path function $L_2^{(i)}(\bm{r})$ is defined as the probability that a
line segment with the length $|\bm{r}|$ and paralel to $\bm{r}$ lies wholly in
the phase $i$ when randomly thrown into the input array. This function for a
one-dimensional array is computed using the algorithm described in
\apref{linpathalg}. A multidimensional arrays is firstly cut in slices as
described above and \code{l2} function is calculated for each slice returning
\code{success} and \code{total} arrays. These arrays are then accumulated as
described in \cref{scansec} for final calculation of $L_2^{(i)}(\bm{r})$
function.

\subsection{Two-point function}
Two-point correlation function $S_2^{(i)}(\bm{r})$ is defined as the probability
that both ends of a line segment with the length $|\bm{r}|$ and paralel to
$\bm{r}$ lie in the phase $i$ when the segment is randomly thrown into the input
array. This can be expressed using the following formula:
\begin{equation*}
  S_2^{(i)}(\bm{r}) = \langle I^{(i)}(\bm{x}) I^{(i)}(\bm{x} + \bm{r}) \rangle
\end{equation*}
where $I^{(i)}(\bm{x})$ is an indicator function for the phase $i$:
\begin{equation*}
I^{(i)}(\bm{x}) = \left\{
\begin{array}{ll}
  1 & \quad \bm{x} \in \text{phase $i$} \\
  0 & \quad \text{otherwise}
\end{array}
\right.
\end{equation*}

Two-point function is so heaviliy used internally in the package so we propose a
bit more general definition for it:
\begin{equation*}
  S_2^{\chi}(\bm{r}) = \langle \chi(\bm{x}, \bm{x} + \bm{r}) \rangle
\end{equation*}
where $\chi(\bm{x})$ is some arbitrary indicator (or membership) function and
$\bm{x} \in \mathbb{R}^{2n}$ where $n$ is dimensionality of the input array. The
vector $\bm{x}$ is constructed as $(\bm{x_1}, \bm{x_2})$ where $\bm{x_1}$ and
$\bm{x_2}$ are the coordinates of the both ends of a line segment thrown into
the input array.

If $\chi(\bm{x})$ can be expressed as
$\chi(\bm{x}) = \chi_1(\bm{x_1})\chi_2(\bm{x_2})$ (what we call
\code{SeparableIndicator} in our package) we use the following algorithm to
calculate the array \code{success} for one-dimensional slice:
\begin{algorithmic}[1]
  \Procedure{s2sep}{$array, \chi_1, \chi_2, periodic$}
  \If{$!periodic$}
    \State $array \gets array \odot zeros(length(array))$ \Comment{$\odot$
      denotes concatenation and $zeros(n)$ returns an array with $n$ zeros.}
  \EndIf
  \State $ind1 \gets \chi_1(array)$ \Comment{apply $\chi_1$ element-wise to $array$}
  \State $ind2 \gets \chi_2(array)$ \Comment{apply $\chi_2$ element-wise to $array$}
  \State $ft1 \gets rfft(ind1)$ \Comment{$rfft$ is fast Fourier transform for
    real-valued data}
  \State $ft2 \gets rfft(ind2)$
  \State $s2ft \gets ft1 \cdot \overline{ft2}$ \Comment{multiply element-wise}
  \State $success \gets irfft(s2ft)$ \Comment{$irfft$ is inverse of $rfft$}
  \State \textbf{return} $success$
  \EndProcedure
\end{algorithmic}

If $\chi(\bm{x})$ cannot be expressed in this way (what we call
\code{InseparableIndicator}) the algorithm is slightly different:
\begin{algorithmic}[1]
  \Procedure{s2insep}{$array, \chi, periodic$}
    \State $len \gets length(array)$
    \If{$periodic$}
      \State $array \gets array \odot array$
    \EndIf
    \State $success \gets zeros(len)$
    \For{$shift \gets 0,len-1$}
      \State $array2 \gets subseq(array, shift)$
      \Comment{$subseq$ returns subsequence of $array$ from the $shift$-th element
        to the end.}
      \State $array[shift] \gets sum(map(\chi, array, array2))$
    \EndFor
    \State \textbf{return} $success$
  \EndProcedure
\end{algorithmic}

@@

The algorithm for calculation of the array \code{success} for
$C_2^{(i)}(\bm{r})$ for one-dimensional slice is straightforward:
\begin{algorithmic}[1]
  \Procedure{c2}{$array, phase, periodic$}
  \State $ind \gets array = phase$ \Comment{$array$ is compared with $phase$
    element-wise}
  \State $clusters \gets clusterize(ind)$
  \Comment{$clusterize(x)$ selects clusters from the array
    $x$. Each cluster gets its own unique integer identificator. The
    identificator $0$ is reserved for the void phase.}
  \State $\chi(x,y) \gets (x = y \ne 0)$
  \State \textbf{return} $s2insep(clusters, \chi, periodic)$
  \EndProcedure
\end{algorithmic}

@@

The algorithm for calculation of the array \code{success} for
$F_{ss}^{(i)}(\bm{r})$ for one-dimensional slice is also straightforward:
\begin{algorithmic}[1]
  \Procedure{surfsurf}{$array, phase, periodic$}
    \State $ind \gets array = phase$ \Comment{$array$ is compared with $phase$
      element-wise}
    \State $edge \gets \text{Detect edges in $ind$}$
    \State \textbf{return} $s2sep(edge, identity, identity, periodic)$
    \Comment{$identity(x) = x$}
  \EndProcedure
\end{algorithmic}

@@

The algorithm for calculation of the array \code{success} for
$F_{sv}^{(i)}(\bm{r})$ for one-dimensional slice is very similar to the previous
one:
\begin{algorithmic}[1]
  \Procedure{surfvoid}{$array, phase, periodic$}
    \State $ind \gets array = phase$ \Comment{$array$ is compared with $phase$
      element-wise}
    \State $edge \gets \text{Detect edges in $ind$}$
    \State $indices \gets \text{Array of all indices in $array$}$
    \State $\chi_1(x) \gets edge[x]$
    \State $\chi_2(x) \gets array[x] = 0$ \Comment{$0$ is a value reserved for the
      void phase}
    \State \textbf{return} $s2sep(indices, \chi_1, \chi_2, periodic)$
  \EndProcedure
\end{algorithmic}

The edges are extracted by calculating $|S(array)|$ where $S$ is the Sobel
operator. \textbf{TODO: Cite our first paper}. Algorithmic efficiency of these
algorithms are $O(n\log n)$ because it's implemented using \code{s2sep} and
filtering to extract edges is $O(n)$.

@@

Our algorithm for calculating $P(r)$ for a two-phase system is the following:
\begin{algorithmic}[1]
  \Procedure{poresize}{$array, nbins$}
    \State $d \gets dist(array)$ \Comment{$dist$ calculates
      $\mathcal{D}(\bm{x})$ for each element of $array$}
    \State $f(x) \gets x \ne 0$
    \State $positive \gets filter(d, f)$
    \State Bin values in $positive$ to $nbins$ bins to compute a histogram $H$
    \State \textbf{return} $H$
  \EndProcedure
\end{algorithmic}
Algorithmic efficiency is $O(n)$ where $n$ is the number of elements in the
input array.

\subsection{Chord length function}
The chord length function $p^{(i)}(\bm{r})d\bm{r}$ is defined as the probability of
finding a chord of length between $|\bm{r}|$ and $|\bm{r} + d\bm{r}|$ and
paralled to $\bm{r}$ in phase $i$. A chords is a line segment between
intersections of an infinitely long line with the interface of the phase
$i$. This function, again, can be calculated by cutting slices from a
multidimensional array. The algorithm which works in $O(n)$ time where $n$ is
the number of elements in the input is as follows:
\begin{algorithmic}[1]
  \Procedure{chordlength}{$array, phase, nbins$}
    \State $lengths \gets [\quad]$
    \State $ind \gets array = phase$ \Comment{$array$ is compared with $phase$
      element-wise}
    \State $edgephase \gets maximum(array) + 1$ \Comment{``bogus'' phase for the
      interface of the phase $phase$}
    \State $edge \gets edgephase \cdot [\text{Detect edges in $ind$}]$ \Comment{We
      use distance transform here}
    \State $combo = min(edge + array, edgephase)$ \Comment{Combine the original
      array and the interface for the phase $phase$}
    \ForAll{$direction \in \{\text{Requested directions}\}$}
      \ForAll{$slice \in \{\text{Slices parallel to $direction$}\}$}
        \State $len \gets 0$
        \State $startonedge \gets false$
        \ForAll{$x \in slice$}
          \If{$x = edgephase$}
            \Comment{Edge found}
            \If{$len > 1 \land startonedge$}
              \State $lengths \gets (len-1):lengths$
            \EndIf
            \State $len \gets 0$
            \State $startonedge \gets true$
          \ElsIf{$x \ne phase$}
            \State $startonedge \gets false$
          \EndIf
          \State $len \gets len + 1$
        \EndFor
      \EndFor
    \EndFor
    \State Bin elements in $lengths$ to $nbins$ bins to compute a histogram $H$
    \State \textbf{return} $H$
  \EndProcedure
\end{algorithmic}

@@

We generate two (resp. three) dimensional square (resp. cubic) binary arrays
with different number of elements. Each element $a_{ij}$ (resp. $a_{ijk}$) of a
two-dimensional (resp. three-dimensional) array is calculated as follows:
\begin{align*}
  a_{ij} & = \left\{
  \begin{array}{ll}
    0 & \quad vn(5i/s, 5j/s, 0, 5, 34365) < 0.5 \\
    1 & \quad \text{otherwise}
  \end{array}
  \right. \\
  a_{ijk} & = \left\{
  \begin{array}{ll}
    0 & \quad vn(5i/s, 5j/s, 5k/s, 5, 34365) < 0.5 \\
    1 & \quad \text{otherwise}
  \end{array}
  \right.
\end{align*}
Here $vn(x,y,z,o,s)$ is a value noise function. We fix a number of
octaves $o = 5$ and a random seed $s = 34365$. The particular implementation for
our measurements was taken from \code{ValueNoise.jl} package. An example of a
two-dimensional image generated in this way is on \cref{fig:valuenoise}.

@@

\appendix
\section{Algorithm of computation of lineal-phase function}
\label{linpathalg}
The algorithm below calculates $L_2^{(phase)}$ correlation function for
one-dimensional array $array$. The result is an array of $L_2$ values for
correlation lengths from $0$ to $maxlen-1$. If $periodic$ is true (resp. false)
then periodic (resp. CW) boundary conditions are used. The entry point of the
algorithm is the function L2. The efficiency of the proposed algorithm is $O(n)$
where $n$ is the length of the array.

\begin{algorithmic}[1]
  \Procedure{countruns}{$array, len, phase$}
    \State $runslist \gets [\quad]$
    \State $runs \gets 0$
    \ForAll{$x \in array$}
      \If{$x = phase$}
        \State $runs \gets runs + 1$
      \ElsIf{$runs \ne 0$}
        \State $runslist \gets runs:runslist$
        \State $runs \gets 0$
      \EndIf
    \EndFor
    \If{$runs \ne 0$}
      \State $runslist \gets runs:runslist$
    \EndIf
    \State \textbf{return} $runslist$
  \EndProcedure
  \\
  \Procedure{updateruns}{$array, runs, n$}
    \For{$idx = 1,n$}
      \State $array[idx] \gets array[idx] + runs$
      \State $runs \gets runs - 1$
    \EndFor
  \EndProcedure
  \\
  \Procedure{updateperiodic}{$array, first, last, n$}
    \State $sum \gets first + last$
    \State $n \gets \min(n, sum)$
    \For{$idx = 1,n$}
      \State $a \gets \min(idx - 1, first, last, sum - (idx - 1))$
      \State $array[idx] \gets array[idx] + a$
    \EndFor
  \EndProcedure
  \\
  \Procedure{l2}{$array, phase, maxlen, periodic$}
    \State $success \gets zeros(maxlen)$
    \State $total \gets zeros(maxlen)$
    \State $len \gets length(array)$
    \State $nupdates \gets \min(len, maxlen)$
    \State $runs \gets countruns(array, maxlen, phase)$
    \ForAll{$run \in runs$}
      \State $updateruns(success, run, \min(run, maxlen))$
    \EndFor
    \If{$periodic$}
      \State $first \gets first(array)$
      \State $last \gets last(array)$
      \If{$first = last$}
        \State $updateperiodic(success, first, last, nupdates)$
      \EndIf
      \State Increment $total[\text{from} \, 1 \, \text{to} \, nupdates]$ by
      $len$
    \Else
      \State $updateruns(total, len, nupdates)$
    \EndIf
    \State \textbf{return} $success$ divided by $total$ element-wise.  
  \EndProcedure
\end{algorithmic}

>>>>

\section{Overview of supported functions and modules.}
\label{oversec}
\verb+CorrelationFunctions.jl+ is capable to compute the following correlation
functions:
\begin{itemize}
\item Lineal-path function $L_2$.
\item Two-point function $S_2$.
\item Cluster function $C_2$.
\item Surface-surface function $F_{ss}$.
\item Surface-void function $F_{sv}$.
\item Pore size function $P$.
\item Chord length function $p$.
\item Phase cross-correlation function $\rho_{ij}$
\end{itemize}

@@

\subsection{Lineal-path function}
Lineal-path function $L_2^{(i)}(\bm{r})$ is defined as the probability that a
line segment with the length $|\bm{r}|$ and paralel to $\bm{r}$ lies wholly in
the phase $i$ when randomly thrown into the input array. This function for a
one-dimensional array is computed using the algorithm described in
\apref{linpathalg}. A multidimensional arrays is firstly cut in slices and
\code{l2} function is calculated for each slice returning \code{success} and
\code{total} arrays. These arrays are then accumulated as described in
\cref{scansec} for final calculation of $L_2^{(i)}(\bm{r})$ function.

\subsection{Two-point function}
\label{sec:s2}
Two-point correlation function $S_2^{(i)}(\bm{r})$ is defined as the probability
that both ends of a line segment with the length $|\bm{r}|$ and paralel to
$\bm{r}$ lie in the phase $i$ when the segment is randomly thrown into the input
array. This can be expressed using the following formula:
\begin{equation*}
  S_2^{(i)}(\bm{r}) = \langle I^{(i)}(\bm{x}) I^{(i)}(\bm{x} + \bm{r}) \rangle
\end{equation*}
where $I^{(i)}(\bm{x})$ is an indicator function for the phase $i$:
\begin{equation*}
I^{(i)}(\bm{x}) = \left\{
\begin{array}{ll}
  1 & \quad \bm{x} \in \text{phase $i$} \\
  0 & \quad \text{otherwise}
\end{array}
\right.
\end{equation*}

Two-point function is so heaviliy used internally in the package so we propose a
bit more general definition for it:
\begin{equation}
  S_2^{\chi}(\bm{r}) = \langle \chi(\bm{x}, \bm{x} + \bm{r}) \rangle
  \label{eq:s2gen}
\end{equation}
where $\chi(\bm{x_1}, \bm{x_2})$ can be understood as some arbitrary indicator
(or membership) function on $\mathbb{R}^{2n}$ where $n$ is dimensionality of the
input array. Vectors $\bm{x_1}$ and $\bm{x_2}$ are the coordinates of the both
ends of a line segment thrown into the input array.

If $\chi(\bm{x_1}, \bm{x_2})$ can be expressed as
$\chi(\bm{x_1}, \bm{x_2}) = \chi_1(\bm{x_1})\chi_2(\bm{x_2})$ (what we call
\code{SeparableIndicator} in our package) we use the following algorithm to
calculate the array \code{success} for one-dimensional slice $a$:
\begin{algorithmic}[1]
  \Procedure{s2sep}{$a, \chi_1, \chi_2, periodic$}
  \If{$!periodic$}
    \State $a \gets a \odot zeros(length(a))$ \Comment{$\odot$
      denotes concatenation and $zeros(n)$ returns an array with $n$ zeros.}
  \EndIf
  \State $a_1 \gets \chi_1(a)$ \Comment{apply $\chi_1$ element-wise to $a$}
  \State $a_2 \gets \chi_2(a)$ \Comment{apply $\chi_2$ element-wise to $a$}
  \State $f_1 \gets rfft(a_1)$ \Comment{$rfft$ is fast Fourier transform for
    real-valued data}
  \State $f_2 \gets rfft(a_2)$
  \State $f_{S_2} \gets f_1 \cdot \overline{f_2}$ \Comment{multiply element-wise}
  \State $success \gets irfft(f_{S_2})$ \Comment{$irfft$ is inverse of $rfft$}
  \State \textbf{return} $success$
  \EndProcedure
\end{algorithmic}

If $\chi(\bm{x})$ cannot be expressed in this way (what we call
\code{InseparableIndicator}) the algorithm is slightly different:
\begin{algorithmic}[1]
  \Procedure{s2insep}{$a, \chi, periodic$}
    \State $len \gets length(a)$
    \If{$periodic$}
      \State $a \gets a \odot a$
    \EndIf
    \State $success \gets zeros(len)$
    \For{$shift \gets 0,len-1$}
      \State $a_2 \gets subseq(a, shift)$
      \Comment{$subseq$ returns subsequence of $a$ from the $shift$-th element
        to the end.}
      \State $success[shift] \gets sum(map(\chi, a, a_2))$
    \EndFor
    \State \textbf{return} $success$
  \EndProcedure
\end{algorithmic}

@@

The algorithm for calculation of the array \code{success} for
$C_2^{(i)}(\bm{r})$ for one-dimensional slice $a$ is straightforward:
\begin{algorithmic}[1]
  \Procedure{c2}{$a, phase, periodic$}
  \State $a_1 \gets x = phase$ \Comment{$a$ is compared with $phase$
    element-wise}
  \State $c \gets clusterize(a_1)$
  \Comment{$clusterize(x)$ selects clusters from the array
    $x$. Each cluster gets its own unique integer identificator. The
    identificator $0$ is reserved for the void phase.}
  \State $\chi(x,y) \gets (x = y \ne 0)$
  \State \textbf{return} $s2insep(c, \chi, periodic)$
  \EndProcedure
\end{algorithmic}

@@

The algorithm for calculation of the array \code{success} for
$F_{ss}^{(i)}(\bm{r})$ for one-dimensional slice $a$is also straightforward:
\begin{algorithmic}[1]
  \Procedure{surfsurf}{$a, phase, periodic$}
    \State $a_1 \gets a = phase$
    \State $edge \gets \text{Detect edges in $a_1$}$
    \State \textbf{return} $s2sep(edge, E, E, periodic)$
    \Comment{$E(x) = x$}
  \EndProcedure
\end{algorithmic}

@@

The algorithm for calculation of the array \code{success} for
$F_{sv}^{(i)}(\bm{r})$ for one-dimensional slice $a$ is very similar to the
previous one:
\begin{algorithmic}[1]
  \Procedure{surfvoid}{$a, phase, periodic$}
    \State $a_1 \gets a = phase$
    \State $edge \gets \text{Detect edges in $a_1$}$
    \State $indices \gets \text{Array of all indices in $a$}$
    \State $\chi_1(x) \gets edge[x]$
    \State $\chi_2(x) \gets a[x] = 0$ \Comment{$0$ is a value reserved for the
      void phase}
    \State \textbf{return} $s2sep(indices, \chi_1, \chi_2, periodic)$
  \EndProcedure
\end{algorithmic}

The edges are extracted by calculating $|S(a)|$ where $S$ is the Sobel
operator. \textbf{TODO: Cite our first paper}. Algorithmic efficiency of these
algorithms are $O(n\log n)$ because it's implemented using \code{s2sep} and
filtering to extract edges is $O(n)$.

@@

Our algorithm for calculating $P(r)$ for a two-phase system $a$ is the
following:
\begin{algorithmic}[1]
  \Procedure{poresize}{$a, nbins$}
    \State $d \gets dist(a)$ \Comment{$dist(a)$ calculates
      $\mathcal{D}(\bm{x})$ for each element of $a$}
    \State $f(x) \gets x \ne 0$
    \State $d' \gets filter(f, d)$
    \State Bin values in $d'$ to $nbins$ bins to compute a histogram $H$
    \State \textbf{return} $H$
  \EndProcedure
\end{algorithmic}
Algorithmic efficiency is $O(n)$ where $n$ is the number of elements in the
input array.

\subsection{Chord length function}
The chord length function $p^{(i)}(\bm{r})d\bm{r}$ is defined as the probability of
finding a chord of length between $|\bm{r}|$ and $|\bm{r} + d\bm{r}|$ and
paralled to $\bm{r}$ in phase $i$. A chord is a line segment between
intersections of an infinitely long line with the interface of the phase
$i$. This function, again, can be calculated by cutting slices from a
multidimensional array. The algorithm which works in $O(n)$ time where $n$ is
the number of elements in the input is as follows:
\begin{algorithmic}[1]
  \Procedure{chordlength}{$a, phase, nbins$}
    \State $lengths \gets [\quad]$
    \State $a_1 \gets a = phase$
    \State $p' \gets maximum(a) + 1$ \Comment{``bogus'' phase for the
      interface of the phase $phase$}
    \State $edge \gets p' \cdot [\text{Detect edges in $a_1$}]$ \Comment{We
      use distance transform here}
    \State $a_2 = min(edge + a, p')$ \Comment{Combine the original
      array and the interface for the phase $phase$}
    \ForAll{$dir \in \{\text{Requested directions}\}$}
      \ForAll{$s \in \{\text{Slices of $a_2$ parallel to $dir$}\}$}
        \State $len \gets 0$
        \State $startonedge \gets false$
        \ForAll{$x \in s$}
          \If{$x = p'$}
            \Comment{Edge found}
            \If{$len > 1 \land startonedge$}
              \State $lengths \gets (len-1):lengths$
            \EndIf
            \State $len \gets 0$
            \State $startonedge \gets true$
          \ElsIf{$x \ne phase$}
            \State $startonedge \gets false$
          \EndIf
          \State $len \gets len + 1$
        \EndFor
      \EndFor
    \EndFor
    \State Bin elements in $lengths$ to $nbins$ bins to compute a histogram $H$
    \State \textbf{return} $H$
  \EndProcedure
\end{algorithmic}

\subsection{Phase cross-corellation function}

Phase cross-correlation function $\rho_{ij}(\bm{r})$ can be defined as
\begin{equation*}
  \rho_{ij}(\bm{r}) = \langle I^{(i)}(\bm{x}) I^{(j)}(\bm{x} + \bm{r}) \rangle
\end{equation*}
which can be rewritten as \cref{eq:s2gen} with $\chi(\bm{x_1}, \bm{x_2})$
defined as follows:
\begin{equation*}
  \chi(\bm{x_1}, \bm{x_2}) = I^{(i)}(\bm{x}) I^{(j)}(\bm{x})
\end{equation*}
Now to calculate cross-correlation function we can use \code{s2sep} function
described in \cref{sec:s2}.

@@

We generate two (resp. three) dimensional square (resp. cubic) binary arrays
with different number of elements. Each element $a_{ij}$ (resp. $a_{ijk}$) of a
two-dimensional (resp. three-dimensional) array is calculated as follows:
\begin{align*}
  a_{ij} & = \left\{
  \begin{array}{ll}
    0 & \quad vn(5i/s, 5j/s, 0, 5, 34365) < 0.5 \\
    1 & \quad \text{otherwise}
  \end{array}
  \right. \\
  a_{ijk} & = \left\{
  \begin{array}{ll}
    0 & \quad vn(5i/s, 5j/s, 5k/s, 5, 34365) < 0.5 \\
    1 & \quad \text{otherwise}
  \end{array}
  \right.
\end{align*}
Here $vn(x,y,z,o,s)$ is a value noise function. We fix a number of
octaves $o = 5$ and a random seed $s = 34365$. The particular implementation of
value noise for our measurements was taken from \code{ValueNoise.jl} package. An
example of a two-dimensional image generated in this way is on
\cref{fig:valuenoise}.

@@

\appendix
\section{Algorithm of computation of lineal-phase function}
\label{linpathalg}
The algorithm below calculates $L_2^{(phase)}$ correlation function for
one-dimensional array $a$. The result is an array of $L_2$ values for
correlation lengths from $0$ to $maxlen-1$. If $periodic$ is true (resp. false)
then periodic (resp. CW) boundary conditions are used. The entry point of the
algorithm is the function L2. The efficiency of the proposed algorithm is $O(n)$
where $n$ is the length of the array.

\begin{algorithmic}[1]
  \Procedure{countruns}{$a, len, phase$}
    \State $runslist \gets [\quad]$
    \State $runs \gets 0$
    \ForAll{$x \in a$}
      \If{$x = phase$}
        \State $runs \gets runs + 1$
      \ElsIf{$runs \ne 0$}
        \State $runslist \gets runs:runslist$
        \State $runs \gets 0$
      \EndIf
    \EndFor
    \If{$runs \ne 0$}
      \State $runslist \gets runs:runslist$
    \EndIf
    \State \textbf{return} $runslist$
  \EndProcedure
  \\
  \Procedure{updateruns}{$a, runs, n$}
    \For{$idx = 1,n$}
      \State $a[idx] \gets a[idx] + runs$
      \State $runs \gets runs - 1$
    \EndFor
  \EndProcedure
  \\
  \Procedure{updateperiodic}{$a, first, last, n$}
    \State $sum \gets first + last$
    \State $n \gets \min(n, sum)$
    \For{$idx = 1,n$}
      \State $a \gets \min(idx - 1, first, last, sum - (idx - 1))$
      \State $a[idx] \gets a[idx] + a$
    \EndFor
  \EndProcedure
  \\
  \Procedure{l2}{$a, phase, maxlen, periodic$}
    \State $success \gets zeros(maxlen)$
    \State $total \gets zeros(maxlen)$
    \State $len \gets length(a)$
    \State $nupdates \gets \min(len, maxlen)$
    \State $runs \gets countruns(a, maxlen, phase)$
    \ForAll{$run \in runs$}
      \State $updateruns(success, run, \min(run, maxlen))$
    \EndFor
    \If{$periodic$}
      \State $first \gets first(a)$
      \State $last \gets last(a)$
      \If{$first = last$}
        \State $updateperiodic(success, first, last, nupdates)$
      \EndIf
      \State Increment $total[\text{from} \, 1 \, \text{to} \, nupdates]$ by
      $len$
    \Else
      \State $updateruns(total, len, nupdates)$
    \EndIf
    \State \textbf{return} $success$ divided by $total$ element-wise.  
  \EndProcedure
\end{algorithmic}
